# Chapter 2: My data analysis

- #2.1 Opening the data and checking the structure. Let's see what happens!

```{r}

learning2014 <- read.table(file="http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", header=T, sep=",")

str(learning2014)

```
- The data has 166 observations and now 7 variables. Variables contain  gender, age, points from the exam and attitude towards statistics. There are three sum variables from combining questions measuring "strategic approach", "deep approach", and "surface approach".

- #2.2 Let's first look at he basic numbers of the data. After that we can begin to analyse it.

```{r}
summary(learning2014)
head(learning2014)
str(learning2014)

```
There is about twice as much females comparered to men. Age ranges from 17 to 55 with mean being 25.51. This gives us a big range in the scale of age. The variable points has a mean of 22.72 with the minimum being 7.00 and the maximum being 33.0.

Now we can also make a graphical overview of the data with ggplot2.

```{r}
# access the GGally and ggplot2 libraries
library(GGally)
library(ggplot2)

p <- ggpairs(learning2014, mapping = aes(col=gender, alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
p
```

The grapical overview could have been a little bit better. I deciced to highlight the gender as I think it was the most important variable in the model. 

From the graph we can see the correlations between the different variables. The highest correlation is between attitude and points. The lowest correlation is a close battle with age - attitude and age - points. Overall the variable age seems to have quite low correlations compared to the others.

- #2.3 Now we fit the model
```{r}
malli <- lm(points ~ attitude + stra + surf, data=learning2014)
summary(malli)
```
The only variable with enough significance in p-value is attitude. This is the reason why we remove stra and surf variables from the model. Now we fit it again.

```{r}
malli <- lm(points ~ attitude, data=learning2014)
summary(malli)
```
Whoo! We got the model!
```{r}
library(ggplot2)
plot(malli, which=c(1, 2,5), col = "turquoise2",  lwd = 3)
```

