#Chapter 4: Clustering and stuff'
-> Lets begin!

We begin this week by analysing data from R's MASS package.

```{r}

library(MASS)
library(corrplot)
library(tidyverse)

data("Boston")

str(Boston)

dim(Boston)

```

The data has 506 observations and 14 different variables. Next we plot the data to show the graphical overview and then we check how different variables compare to each other.

```{r}
summary(Boston)

```
Tähän kommentti yleisesti datasta.


```{r}

cor_matrix<-cor(Boston) 


cor_matrix %>% round(digits = 2)
wb <- c("white","black") 

corrplot(cor_matrix, order="hclust", addrect=2, bg="gold2", col=wb)
 

```
Tähän kommentti muuttujien suhteesta toisiinsa.



```{r}

myboston_scaled <- scale(Boston)
summary(myboston_scaled)


```

After we standardized all the means are now 0.


```{r}
cutoffs <- quantile(myboston_scaled$crim)
labels <- c("low", "mod_low", "mod_high", "high")
crim_cat <- cut(myboston_scaled$crim, breaks=cutoffs, include.lowest = TRUE, label = labels)
table(crim_cat)

```



```{r}
n <- nrow(myboston_scaled)
eighty_perc <- sample(n, size = n * 0.8)
train <- myboston_scaled[eighty_perc,]
test <- myboston_scaled[-eighty_perc,]

```


