#Chapter 5: Dimensionality reduction techniques

Lets begin the Chapter 5 by loading the data.
```{r, include=FALSE}
library(ggplot2)
library(GGally)
library(MASS)
library(corrplot)
library(tidyverse)
library(plotly)
library(dplyr)
library(FactoMineR)
require(ggplot2)

```

This weeks data contains a large variety of countries. We look for example their life expectancy, GNI and education expectancy.

```{r}
human <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt", header = T, sep=",")

dim(human)
str(human)

```

We have 155 observations (countries) and 8 different variables.
Next we look at the graphical overview of the data and try to see correlations between the variables.


```{r}
p <- ggpairs(human, mapping = aes(col="blue", alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))

p
```
There is a lot of significant correlations between the variables. For example Mat.mor - Ado.birth has 0.759 correlation.

```{r}
summary(human)
```

Next I chose a couple of interesting variables and tried to explain them with GNI also known as Gross National Income ($). I chose two variables: Life Expectancy (in years) and Education Expectancy (in years). 

```{r}
p2 <- ggplot(human, aes(x=GNI, y= Life.Exp)) + geom_point(col="deepskyblue1") + geom_smooth( col = "red2")

p2
```

```{r}

model5 <- lm(Life.Exp ~ GNI, human)
summary(model5)
plot(model5,  which=c(1), col = "turquoise2",  lwd = 3)
```
As we see there are quite a few outliers such as Qatar that make the model hard to fit.

```{r}
p3 <- ggplot(human, aes(x=GNI, y= Edu.Exp)) + geom_point(col="red2") + geom_smooth(col = "royalblue1")

p3
```

Next we perform principal component analysis (PCA) on the not standardized human data. 

```{r}
model6 <- lm(Edu.Exp ~ GNI, human)
summary(model6)
plot(model6,  which=c(1), col = "turquoise2",  lwd = 3)
```
This model also suffers a bit as there are those large outliers such as Niger and Qatar.

```{r, warning=FALSE}
# perform principal component analysis (with the SVD method)
pca_human <- prcomp(human)

# draw a biplot of the principal component representation and the original variables
biplot(pca_human, choices = 1:2, cex = c(0.5, 1), col = c("black", "red2"), main="Biplot of the first two principal components for the unscaled data")

```

The picture doesn't look very good as we have not scaled the data. This is the reason why it is important to scale the data.


```{r}
human_std <- scale(human)

pca_human11 <- prcomp(human_std)

summary(pca_human11)

```


```{r}

biplot(pca_human11, choices = 1:2, cex = c(0.6, 1.2), col = c("black", "red2"),main="Biplot of the first two principal components for the scaled data")


```


Next we load the Tea dataset from Factominer package.
```{r}

data("tea")
dim(tea)
str(tea)

```


I decided to drop out some variables as there was 36 of them. It was quite impossible to plot and make a grapical overview with a such variety of different variables.
```{r}
keep <- c("friends", "Tea", "resto", "age_Q", "frequency")

teatimes <- dplyr::select(tea, one_of(keep))

str(teatimes)


```


```{r, include= FALSE}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

```{r}
g1 <- ggplot(teatimes, aes(friends)) + geom_bar() + theme_grey() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))
g2 <- ggplot(teatimes, aes(Tea)) + geom_bar()+ theme_grey() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))
g3 <- ggplot(teatimes, aes(resto)) + geom_bar() + theme_grey() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))
g4 <- ggplot(teatimes, aes(age_Q)) + geom_bar()+ theme_grey() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))
g5 <- ggplot(teatimes, aes(frequency)) + geom_bar()+ theme_grey() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 10))

multiplot(g1, g2, g3, g4, g5, cols=3)


```


```{r}
mca <- MCA(X = teatimes, graph=F)

summary(mca)


```


```{r}
plot(mca, invisible=c("ind"), habillage = "quali")

```

