# Chapter 2: My data analysis

- #2.1 Opening the data and checking the structure. Let's see what happens!

```{r}

learning2014 <- read.table(file="http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt", header=T, sep=",")

str(learning2014)

```
- The data has 166 observations and now 7 variables. Variables contain gender, age, attitude, deep learning, strategic learning, surface learning and points.

- Then lets draw the data and see how it look graphically.

```{r}
# access the GGally and ggplot2 libraries
library(GGally)
library(ggplot2)

p <- ggpairs(learning2014, mapping = aes(col=gender, alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
p
summary(learning2014)
```

```{r}
malli <- lm(points ~ attitude + stra + surf, data=learning2014)
summary(malli)
```
The only variable with enough significance in p-value is attitude. This is the reason why we remove stra and surf variables from the model.

```{r}
malli <- lm(points ~ attitude, data=learning2014)
summary(malli)
```
Whoo! We got the model!
```{r}
library(ggplot2)
plot(malli, which=c(1, 2,5), col = "turquoise2",  lwd = 3)
```

