# Chapter 3: My data analysis on week 3

```{r} 
alc <- read.table("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/alc.txt ", header=T, sep=",")

colnames(alc)

```
# The data has 382 observations and 35 variables after we did the following changes:
 - The variables not used for joining the two data have been combined by averaging  (including the grade variables)
 - 'alc_use' is the average of 'Dalc' and 'Walc'
 - 'high_use' is TRUE if 'alc_use' is higher than 2 and FALSE otherwise

The data is about students alcohol consumption. 

I'm going to conduct an analysis on relationships between high/low alcohol consumption and sex, age and the grades G1, G2 and G3.

My hypothesis is that high alcohol consumption leads to lower grades. Also I want to see if age or sex has some kind of an effect.

# Drawing the plots and figuring out how my hypothesis begins to look.

```{r}
library(ggplot2)
g1 <- ggplot(alc, aes(x = high_use, y = G1, col = sex))
g1 + geom_boxplot() + ylab("grade")

g2 <- ggplot(alc, aes(x = high_use, y = G2, col = sex))
g2 + geom_boxplot() + ylab("grade")

g3 <- ggplot(alc, aes(x = high_use, y = G3, col = sex))
g3 + geom_boxplot() + ylab("grade")

```